Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2023-04-12 src/tex/ms.tex
Squashed commit of the following:
commit ce82d71beb71eaa0e354018553cc13aa6f14c49a
Author: Kaze Wong <kazewong.physics@gmail.com>
Date:   Fri Mar 31 15:23:56 2023 -0400

    Chat

commit 71f0fc4d653ad2396ebcf10be965fca3d70ff24d
Author: Kaze Wong <kazewong.physics@gmail.com>
Date:   Fri Mar 31 11:59:08 2023 -0400

    GPT pass

commit b18cc5b5dcc5fd80135536bcbc414c6337b98577
Author: Kaze Wong <kazewong.physics@gmail.com>
Date:   Fri Mar 31 11:44:24 2023 -0400

    GPT pass of Result

Given the ansatz used in the waveform model is unlikely to be fully compatible
with NR, and the optimization procedure is done over a distribution of waveforms
with different source parameters, it is conceivable that there are some
trade-off in accuracy of the waveform models between different part of the
parameter space. If this is the reason why the high mismatch tail is not
improving during the joint-optimization, separating the parameter space into
smaller subspace should help alleviate this issue. On the other hand, if the
ansatz does not have the right parameterized form to capture the behavior of the
NR waveforms as a function of the intrinsic parameters, the result should be
always biased, and we should not expect any improvement even if we separate the
parameter space into smaller space during training.

Since we know intrinsic parameters play an important role in the ansatz, we
would like to investigate how intrinsic parameters affect the recalibration
process. First, we plot the parameter space of $q$ vs. $\chi_{\mathrm{PN}}$ in
Fig. \ref{fig:ps_q148_qchi}. In the low mass ratio region, waveforms with both
positive and negative log differences are mixed up. On the other hand, along the
horizontal line $\chi_{\mathrm{PN}}=0$, waveform mismatches consistently improve
under recalibration.

Furthermore, we plot the parameter space of $\chi_1$ vs. $\chi_2$ in
Fig.~\ref{fig:ps_q148}. Waveforms along the diagonal axis, i.e.
$\chi_1\approx\chi_2$, show good mismatch improvements. We expect to see this
feature since the original coefficients were fitted using NR waveforms with
equal or similar spin, hence the model prefers waveforms with similar spin. One
interesting feature is how the second and fourth quadrants respond to
optimization. In the second quadrant ($\chi_1<0$ and $\chi_2>0$), waveforms
generally improve with along optimization. However, mismatches in the fourth
quadrant ($\chi_1>0$ and $\chi_2<0$) do not improve after optimization. Most
waveforms even turned worse after optimization. These waveforms correspond to
the waveforms in the high mismatch tail in Fig.~\ref{fig:q148}.

% Waveform mismatches in the fourth quadrant, i.e. $\chi_1>0$ and $\chi_2<0$,
% performed worse after optimized. IMRPhenomD is constructed using single spin
% parameter, where there are degeneracy between $\chi_1$ and $\chi_2$. In
% regions with opposite extreme spin components, degeneracy significantly alters
% intrinsic parameters passed into waveform generation, hence the fourth
% quadrant shows such feature.

>>>>

Given that the waveform model's ansatz may not be entirely compatible with NR,
and the optimization procedure is carried out over a distribution of waveforms
with varying source parameters, it is conceivable that some trade-offs in
accuracy exist between different parts of the parameter space. If this is the
cause of the high mismatch tail's failure to improve during joint-optimization,
segmenting the parameter space into smaller subspaces should alleviate this
problem. On the other hand, if the ansatz lacks the correct parameterized form
to capture the NR waveforms' behavior as a function of the intrinsic parameters,
the results will always be biased, and we should not expect any improvement,
even if we segment the parameter space during training.

Since we know intrinsic parameters play an important role in the ansatz, we
would like to investigate how intrinsic parameters affect the recalibration
process. First, we plot the parameter space of $q$ vs. $\chi_{\mathrm{PN}}$ in
Fig. \ref{fig:ps_q148_qchi}. We can see near non-spinning waveforms demonstrate
more consistent improvement, whereas spinning waveforms can either improve or
worsen in terms of mismatch. \kw{What does this means?}

Furthermore, we plot the parameter space of $\chi_1$ vs. $\chi_2$ in
Fig.~\ref{fig:ps_q148}. Waveforms along the diagonal axis, i.e.
$\chi_1\approx\chi_2$, show good mismatch improvements. We expect to see this
feature since the original coefficients were fitted using NR waveforms with
equal or similar spin, hence the model prefers waveforms with similar spin. One
interesting feature is how the second and fourth quadrants respond to
optimization. \kw{How did you count quadrant?} In the second quadrant ($\chi_1<0$ and $\chi_2>0$), waveforms
generally improve with along optimization. However, mismatches in the fourth
quadrant ($\chi_1>0$ and $\chi_2<0$) do not improve after optimization. Most
waveforms even turned worse after optimization. These waveforms correspond to
the waveforms in the high mismatch tail in Fig.~\ref{fig:q148}. \kw{Give justification.}
